How cryptic is cryptic diversity? Machine learning approaches to classifying morphological variation in the Pacific Pond Turtle (\textit{Emys marmorata})

### 2015-09-10 doc/draft_turtle.tex
submitted to syst biol

\begin{abstract}
  We investigate the morphometric identification of cryptic species using machine-learning approaches by examining their implications for a recently proposed cryptic turtle species (\textit{Emys pallida}). We collected landmark-based morphometric data from 354 adult \textit{E. marmorata/ ``pallida''} museum specimens. We assigned a classification to each specimen for six different binning schemes based on geographic occurrence data recorded in museum collection archives. We used multiple machine learning methods, both unsupervised and supervised, to compare different classification hypotheses and asked whether it is possible to determine which amongst a set of classification hypotheses is best. In addition, we applied the above approach to a “clear-cut” example of seven unambiguously distinct species closely related to \textit{E. marmorata}. The results of this study indicate that there is no clear grouping of \textit{E. marmorata/ ``pallida''} based on plastron shape. In contrast, the analysis of the ``clear-cut'' example demonstrates a near perfect classification, which demonstrates that the methods can recover correct results when an appropriate signal exists. Explanations for the lack of grouping in \textit{E. marmorata} include that possibility that genetic differentiation is not associated with plastron shape variation below the species level and/or that local selective pressures (e.g., from hydrological regime) overwhelm morphological differentiation. A reconsideration of the methods used to delimit \textit{E. ``pallida,''} the lack of barriers to gene flow, the strong evidence for widespread admixture between lineages, and the fact that plastron shape can be used to differentiate other emydine species suggest that its lack of diagnosability most likely reflects the non-distinctiveness of this proposed taxon.
\end{abstract}

@@

Here, we investigate the morphometric identification of cryptic species using machine-learning approaches. In particular, we ask whether it is possible to determine which amongst a set of classification hypotheses is best and examine the implications of the results for a recently proposed set of cryptic turtle species.

@@

\clearpage
\bibliographystyle{sysbio}
\bibliography{turtle,packages}

@@

\begin{table}[ht]
  \centering
  \include{comp_tab}
  \caption{AUC values for the best model of each classification scheme for both the observed (training) data and the generalized (testing) data. Results from all three different supervised learning approaches are shown here. AUC values range between 0.5 and 1. }
  \label{tab:comp}
\end{table}

\begin{table}
  \centering
  \begin{tabular}{ l r r r }
    \hline
    Scheme & RF & MLR & LDA \\
    \hline
    \hline
    Morph 1 & \(\ast\) & \(\ast\) & 1 \\
    Morph 2 & 0.79 & 0.55 & 1 \\
    Mito 1 & 0.89 & 0.94 & 1 \\ 
    Mito 2 & 0.82 & 0.57 & \(\ast\) \\ 
    Mito 3 & 0.79 & 0.69 & 0.73 \\ 
    Nuclear & 0.79 & 0.96 & 0.96 \\ 
    \hline
  \end{tabular}
  \caption{Results of bootstrap comparisons between the scheme with the highest mean AUC value and all other schemes, defined P(best - other \(>\) 0). An asterix indicates the best scheme. This was done for each of the three modeling techniques included in this study: random forest (RF), multinomial logistic regression (MLR), and linear discriminate analysis (LDA). Probabilities are the percent of comparisons that are greater than the observed difference in means.}
  \label{tab:gen_tests}
\end{table}

>>>>

\begin{abstract}
  We investigate the morphometric identification of cryptic species using machine learning approaches by examining their implications for a recently proposed cryptic turtle species (\textit{Emys pallida}). We collected landmark-based morphometric data from 354 adult \textit{E. marmorata/``pallida''} museum specimens. We assigned a classification to each specimen for six different binning schemes based on geographic occurrence data recorded in museum collection archives. We used multiple machine learning methods, both unsupervised and supervised, to compare different classification hypotheses and asked whether it is possible to determine which amongst a set of classification hypotheses is best. In addition, we applied the above approach to a “clear-cut” example of seven unambiguously distinct species closely related to \textit{E. marmorata}. The results of this study indicate that there is no clear grouping of \textit{E. marmorata/``pallida''} based on plastron shape. In contrast, the analysis of the ``clear-cut'' example demonstrates a near perfect classification, which demonstrates that the methods can recover correct results when an appropriate signal exists. Explanations for the lack of grouping in \textit{E. marmorata} include that possibility that genetic differentiation is not associated with plastron shape variation below the species level and/or that local selective pressures (e.g., from hydrological regime) overwhelm morphological differentiation. A reconsideration of the methods used to delimit \textit{E. ``pallida,''} the lack of barriers to gene flow, the strong evidence for widespread admixture between lineages, and the fact that plastron shape can be used to differentiate other emydine species suggest that its lack of diagnosability most likely reflects the non-distinctiveness of this proposed taxon.
\end{abstract}

@@

Here, we investigate the morphometric identification of cryptic species using machine learning approaches. In particular, we ask whether it is possible to determine which amongst a set of classification hypotheses is best and examine the implications of the results for a recently proposed set of cryptic turtle species.

@@

\clearpage
\begin{thebibliography}{77}
\providecommand{\natexlab}[1]{#1}
\providecommand{\selectlanguage}[1]{\relax}
\providecommand{\bibAnnoteFile}[1]{%
  \IfFileExists{#1}{\begin{quotation}\noindent\textsc{Key:} #1\\
  \textsc{Annotation:}\ \input{#1}\end{quotation}}{}}
\providecommand{\bibAnnote}[2]{%
  \begin{quotation}\noindent\textsc{Key:} #1\\
  \textsc{Annotation:}\ #2\end{quotation}}

\bibitem[{Angielczyk and Feldman(2013)}]{Angielczyk2013a}
Angielczyk, K.~D. and C.~R. Feldman. 2013. {Are diminutive turtles
  miniaturized? The ontogeny of plastron shape in emydine turtles}. Biological
  Journal of the Linnean Society 108:727--755.
\bibAnnoteFile{Angielczyk2013a}

\bibitem[{Angielczyk et~al.(2011)Angielczyk, Feldman, and
  Miller}]{Angielczyk2011}
Angielczyk, K.~D., C.~R. Feldman, and G.~R. Miller. 2011. {Adaptive evolution
  of plastron shape in emydine turtles.} Evolution 65:377--394.
\bibAnnoteFile{Angielczyk2011}

\bibitem[{Angielczyk and Sheets(2007)}]{Angielczyk2007}
Angielczyk, K.~D. and H.~D. Sheets. 2007. {Investigation of simulated tectonic
  deformation in fossils using geometric morphometrics}. Paleobiology
  33:125--148.
\bibAnnoteFile{Angielczyk2007}

\bibitem[{Baird and Girard(1852)}]{Baird1852}
Baird, S.~F. and C.~Girard. 1852. {Descriptions of new species of reptiles
  collected by the U.S. Exploring Expedition under the command of Capt. Charles
  Wilkes}. Proceedings of the National Academy of Sciences Philadelphia
  6:174--177.
\bibAnnoteFile{Baird1852}

\bibitem[{Bauer et~al.(2000)Bauer, Parham, Brown, Stuart, Grismer, Papenfuss,
  Bohme, Savage, Carranza, Grismer, Wagner, Schmitz, Ananjeva, and
  Inger}]{Bauer2000}
Bauer, A.~M., J.~F. Parham, R.~M. Brown, B.~L. Stuart, L.~Grismer, T.~J.
  Papenfuss, W.~Bohme, J.~M. Savage, S.~Carranza, J.~L. Grismer, P.~Wagner,
  A.~Schmitz, N.~B. Ananjeva, and R.~F. Inger. 2000. {Availability of new
  Baysian-delimited gecko names and the importance of character-based species
  descriptions}. Proceedings of the Royal Society B: Biological Sciences
  278:490--492.
\bibAnnoteFile{Bauer2000}

\bibitem[{Baylac et~al.(2003)Baylac, Villemant, and Simbolotti}]{Baylac2003}
Baylac, M., C.~Villemant, and G.~Simbolotti. 2003. {Combining geometric
  morphometrics with pattern recognition for the investigation of species
  complexes}. Biological Journal of the Linnean Society 80:89--98.
\bibAnnoteFile{Baylac2003}

\bibitem[{Bickford et~al.(2007)Bickford, Lohman, Sodhi, Ng, Meier, Winker,
  Ingram, and Das}]{Bickford2007}
Bickford, D., D.~J. Lohman, N.~S. Sodhi, P.~K.~L. Ng, R.~Meier, K.~Winker,
  K.~K. Ingram, and I.~Das. 2007. {Cryptic species as a window on diversity and
  conservation.} Trends in ecology \& evolution 22:148--55.
\bibAnnoteFile{Bickford2007}

\bibitem[{Breiman(2001)}]{Breiman2001}
Breiman, L. 2001. {Random Forests}. Machine Learning 45:5--32.
\bibAnnoteFile{Breiman2001}

\bibitem[{Breiman et~al.(1984)Breiman, Friedman, Stone, and
  Olshen}]{Breiman1984}
Breiman, L., J.~Friedman, C.~J. Stone, and R.~A. Olshen. 1984. {Classification
  and regression trees}. Wadsworth International Group, Belmont.
\bibAnnoteFile{Breiman1984}

\bibitem[{Bury et~al.(2010)Bury, Germano, and Bury}]{Bury2010}
Bury, R.~B., D.~J. Germano, and G.~W. Bury. 2010. {Population Structure and
  Growth of the Turtle Actinemys marmorata from the Klamath–Siskiyou
  Ecoregion: Age, Not Size, Matters}. Copeia 2010:443--451.
\bibAnnoteFile{Bury2010}

\bibitem[{Buskirk et~al.(2005)Buskirk, Parham, and Feldman}]{Buskirk2005}
Buskirk, S.~W., J.~F. Parham, and C.~R. Feldman. 2005. {On the hybridisation
  between two distantly related Asian turtles (Testudines: Scalia x Mauremys)}.
  Salamandra 41:21--26.
\bibAnnoteFile{Buskirk2005}

\bibitem[{Cardini et~al.(2009)Cardini, Nagorsen, O'Higgins, Polly, {Thorington
  Jr}, and Tongiorgi}]{Cardini2009a}
Cardini, A., D.~Nagorsen, P.~O'Higgins, P.~D. Polly, R.~W. {Thorington Jr}, and
  P.~Tongiorgi. 2009. {Detecting biological distinctiveness using geometric
  morphometrics: an example case from the Vancouver Island marmot}. Ethology
  Ecology \& Evolution 21:209--223.
\bibAnnoteFile{Cardini2009a}

\bibitem[{Carstens and Dewey(2010)}]{Carstens2010}
Carstens, B.~C. and T.~A. Dewey. 2010. {Species Delimitation Using a Combined
  Coalescent and Information-Theoretic Approach: An Example from North American
  Myotis Bats}. Systematic Biology 59:400--414.
\bibAnnoteFile{Carstens2010}

\bibitem[{Carstens et~al.(2013)Carstens, Pelletier, Reid, and
  Satler}]{Carstens2013}
Carstens, B.~C., T.~a. Pelletier, N.~M. Reid, and J.~D. Satler. 2013. {How to
  fail at species delimitation.} Molecular ecology 22:4369--83.
\bibAnnoteFile{Carstens2013}

\bibitem[{Caumul and Polly(2005)}]{Caumul2005a}
Caumul, R. and P.~D. Polly. 2005. {Phylogenetic and environmental components of
  morphological variation: skull, mandible, and molar shape in marmots
  (Marmota, Rodentia).} Evolution; international journal of organic evolution
  59:2460--72.
\bibAnnoteFile{Caumul2005a}

\bibitem[{Clare(2011)}]{Clare2011}
Clare, E.~L. 2011. {Cryptic species? Patterns of maternal and paternal gene
  flow in eight neotropical bats.} PloS one 6:e21460.
\bibAnnoteFile{Clare2011}

\bibitem[{Dayrat(2005)}]{Dayrat2005a}
Dayrat, B. 2005. {Towards integrative taxonomy}. Biological Journal of the
  Linnean Society 85:407--415.
\bibAnnoteFile{Dayrat2005a}

\bibitem[{{De Queiroz}(2007)}]{DeQueiroz2007b}
{De Queiroz}, K. 2007. {Species concepts and species delimitation.} Systematic
  Biology 56:879--86.
\bibAnnoteFile{DeQueiroz2007b}

\bibitem[{Demandt and Bergek(2009)}]{Demandt2009}
Demandt, M.~H. and S.~Bergek. 2009. {Identification of cyprinid hybrids by
  using geometric morphometrics and microsatellites}. Journal of Applied
  Ichthyology 25:695--701.
\bibAnnoteFile{Demandt2009}

\bibitem[{Dobigny et~al.(2003)Dobigny, Granjon, Aniskin, Ba, and
  Voloboulev}]{Dobigny2003}
Dobigny, G., L.~Granjon, V.~Aniskin, K.~Ba, and V.~Voloboulev. 2003. {A new
  sigling species of Taterillus (Muridae, Gerbillinae) from West Agrica}.
  Mammalian Biology 68:299--316.
\bibAnnoteFile{Dobigny2003}

\bibitem[{Dryden(2013)}]{Dryden2013}
Dryden, I.~L. 2013. shapes: Statistical shape analysis. R package version
  1.1-8.
\bibAnnoteFile{Dryden2013}

\bibitem[{Dryden and Mardia(1998)}]{Dryden1998a}
Dryden, I.~L. and K.~Y. Mardia. 1998. {Statistical shape analysis}. Wiley, New
  York.
\bibAnnoteFile{Dryden1998a}

\bibitem[{Eldredge and Gould(1972)}]{Eldredge1972}
Eldredge, N. and S.~J. Gould. 1972. {Punctuated equilibria: an alternative to
  phyletic gradualism}. Pages~82--115 \emph{in} Models in Paleobiology
  (T.~J.~M. Schopf, ed.). Freeman Cooper, San Francisco.
\bibAnnoteFile{Eldredge1972}

\bibitem[{Feldman and Parham(2002)}]{Feldman2002}
Feldman, C.~R. and J.~F. Parham. 2002. {Molecular phylogenetics of emydine
  turtles: taxonomic revision and the evolution of shell kinesis.} Molecular
  Phylogenetics and Evolution 22:388--98.
\bibAnnoteFile{Feldman2002}

\bibitem[{Francoy et~al.(2009)Francoy, Silva, Nunes-Silva, Menezes, and
  Imperatriz-Fonseca}]{Francoy2009}
Francoy, T.~M., R.~A.~O. Silva, P.~Nunes-Silva, C.~Menezes, and V.~L.
  Imperatriz-Fonseca. 2009. {Gender identification of five genera of stingless
  bees (Apidae, Meliponini) based on wing morphology}. Genetics and molecular
  research 8:207--214.
\bibAnnoteFile{Francoy2009}

\bibitem[{Funk et~al.(2012)Funk, Caminer, and Ron}]{Funk2012}
Funk, W.~C., M.~Caminer, and S.~R. Ron. 2012. {High levels of cryptic species
  diversity uncovered in Amazonian frogs.} Proceedings of the Royal Society B:
  Biological Sciences 279:1806--14.
\bibAnnoteFile{Funk2012}

\bibitem[{Gaubert et~al.(2005)Gaubert, Taylor, Fernandes, Bruford, and
  Veron}]{Gaubert2005b}
Gaubert, P., P.~J. Taylor, C.~a. Fernandes, M.~W. Bruford, and G.~Veron. 2005.
  {Patterns of cryptic hybridization revealed using an integrative approach: a
  case study on genets (Carnivora, Viverridae, Genetta spp.) from the southern
  African subregion}. Biological Journal of the Linnean Society 86:11--33.
\bibAnnoteFile{Gaubert2005b}

\bibitem[{Germano and Bury(2009)}]{Germano2009}
Germano, D.~J. and R.~B. Bury. 2009. {Variation in body size, growth, and
  population structure of Actinemys marmorata from lentic and lotic habitats in
  Southern Oregon}. Journal of Herpetology 43:510--520.
\bibAnnoteFile{Germano2009}

\bibitem[{Germano and Rathbun(2008)}]{Germano2008}
Germano, D.~J. and G.~B. Rathbun. 2008. {Growth, population structure, and
  reproduction of western pond turtles (Actinemys marmorata) on the Central
  Coast of California}. Chelonian Conservation and Biology 7:188--194.
\bibAnnoteFile{Germano2008}

\bibitem[{Gould and Eldredge(1977)}]{Gould1977a}
Gould, S.~J. and N.~Eldredge. 1977. {Punctuated equilibria: the tempo and mode
  of evolution reconsidered}. Paleobiology 3:115--151.
\bibAnnoteFile{Gould1977a}

\bibitem[{G\"{u}nd\"{u}z et~al.(2007)G\"{u}nd\"{u}z, Jaarola, Tez, Yeniyurt,
  Polly, and Searle}]{Gunduz2007}
G\"{u}nd\"{u}z, I., M.~Jaarola, C.~Tez, C.~Yeniyurt, P.~D. Polly, and J.~B.
  Searle. 2007. {Multigenic and morphometric differentiation of ground
  squirrels (Spermophilus, Scuiridae, Rodentia) in Turkey, with a description
  of a new species.} Molecular phylogenetics and evolution 43:916--35.
\bibAnnoteFile{Gunduz2007}

\bibitem[{Hand and Till(2001)}]{Hand2001}
Hand, D.~J. and R.~J. Till. 2001. {A Simple Generalisation of the Area Under
  the ROC Curve for Multiple Class Classification Problems}. Machine Learning
  45:171--186.
\bibAnnoteFile{Hand2001}

\bibitem[{Hastie et~al.(2009)Hastie, Tibshirani, and Friedman}]{Hastie2009}
Hastie, T., R.~Tibshirani, and J.~Friedman. 2009. {The elements of statistical
  learning: data mining, inference, and prediction}. 2nd ed. Springer, New
  York.
\bibAnnoteFile{Hastie2009}

\bibitem[{Hausdorf and Hennig(2010)}]{Hausdorf2010}
Hausdorf, B. and C.~Hennig. 2010. {Species delimitation using dominant and
  codominant multilocus markers.} Systematic biology 59:491--503.
\bibAnnoteFile{Hausdorf2010}

\bibitem[{Holland(1992)}]{Holland1992}
Holland, D.~C. 1992. {Level and pattern in morphological variation: a
  phylogeographic study of the western pond turtle (Clemmys marmorata)}. Ph.D.
  thesis University of Southwestern Louisiana.
\bibAnnoteFile{Holland1992}

\bibitem[{Huelsenbeck et~al.(2011)Huelsenbeck, Andolfatto, and
  Huelsenbeck}]{Huelsenbeck2011b}
Huelsenbeck, J.~P., P.~Andolfatto, and E.~T. Huelsenbeck. 2011. {Structurama:
  bayesian inference of population structure.} Evolutionary bioinformatics
  online 7:55--9.
\bibAnnoteFile{Huelsenbeck2011b}

\bibitem[{Hunt(2008)}]{Hunt2008}
Hunt, G. 2008. {Gradual or pulsed evolution: when should punctuational
  explanations be preferred?} Paleobiology 34:360--377.
\bibAnnoteFile{Hunt2008}

\bibitem[{Kaufman and Rousseeuw(1990)}]{Kaufman1990}
Kaufman, L. and P.~J. Rousseeuw. 1990. {Finding groups in data : an
  introduction to cluster analysis}. Wiley, New York.
\bibAnnoteFile{Kaufman1990}

\bibitem[{Kendall(1977)}]{Kendall1977a}
Kendall, D.~G. 1977. {The diffusion of shape}. Advances in Applied Probability
  9:428--430.
\bibAnnoteFile{Kendall1977a}

\bibitem[{Kendall(1984)}]{Kendall1984a}
Kendall, D.~G. 1984. {Shape Manifolds, Procrustean Metrics, and Complex
  Projective Spaces}. Bulletin of the London Mathematical Society 16:81--121.
\bibAnnoteFile{Kendall1984a}

\bibitem[{Klingenberg et~al.(2002)Klingenberg, Barluenga, and
  Meyer}]{Klingenberg2002}
Klingenberg, C.~P., M.~Barluenga, and A.~Meyer. 2002. {Shape analysis of
  symetric structures: quantifying variation among individuals and asymmetry}.
  Evolution 56:1909--1920.
\bibAnnoteFile{Klingenberg2002}

\bibitem[{Kuhn(2013)}]{KuhnMAN2013}
Kuhn, M. 2013. caret: Classification and Regression Training. R package version
  5.15-61.
\bibAnnoteFile{KuhnMAN2013}

\bibitem[{Kuhn and Johnson(2013)}]{Kuhn2013}
Kuhn, M. and K.~Johnson. 2013. {Applied predictive modeling}. Springer, New
  York, NY.
\bibAnnoteFile{Kuhn2013}

\bibitem[{Leach\'{e} and Fujita(2010)}]{Leache2010}
Leach\'{e}, A.~D. and M.~K. Fujita. 2010. {Bayesian species delimitation in
  West African forest geckos (Hemidactylus fasciatus).} Proceedings. Biological
  sciences / The Royal Society 277:3071--7.
\bibAnnoteFile{Leache2010}

\bibitem[{Lubcke and Wilson(2007)}]{Lubcke2007}
Lubcke, G.~M. and D.~S. Wilson. 2007. {Variation in shell morphology of the
  Western Pond Turtle (Actinemys marmorata Baird and Giarard) from three
  aquativ habitats in Northern California}. Journal of Herpetology 41:107--114.
\bibAnnoteFile{Lubcke2007}

\bibitem[{MacLeod(2007)}]{MacLeod2007}
MacLeod, N. 2007. {Automated taxon identification in systematics: theory,
  approaches and applications}. CRC Press, Boca Raton.
\bibAnnoteFile{MacLeod2007}

\bibitem[{Maechler et~al.(2013)Maechler, Rousseeuw, Struyf, Hubert, and
  Hornik}]{Maechler2013}
Maechler, M., P.~Rousseeuw, A.~Struyf, M.~Hubert, and K.~Hornik. 2013. cluster:
  Cluster Analysis Basics and Extensions. R package version 1.14.4.
\bibAnnoteFile{Maechler2013}

\bibitem[{Mitrovski-Bogdanovic et~al.(2013)Mitrovski-Bogdanovic, Petrovic,
  Mitrovic, Ivanovic, \v{Z}ikic, Star\'{y}, Vorburger, and
  Tomanovic}]{MitrovskiBogdanovic2013}
Mitrovski-Bogdanovic, A., A.~Petrovic, M.~Mitrovic, A.~Ivanovic, V.~\v{Z}ikic,
  P.~Star\'{y}, C.~Vorburger, and v.~Tomanovic. 2013. {Identification of two
  cryptic species within the Praon abjectum group (Hymenoptera: Braconidae:
  Aphidiinae) using molecular markers and geometric morphometrics}. Annals of
  the entomological society of America 106:170--180.
\bibAnnoteFile{MitrovskiBogdanovic2013}

\bibitem[{Mitteroecker and Bookstein(2011)}]{Mitteroecker2011}
Mitteroecker, P. and F.~Bookstein. 2011. {Linear Discrimination, Ordination,
  and the Visualization of Selection Gradients in Modern Morphometrics}.
  Evolutionary Biology 38:100--114.
\bibAnnoteFile{Mitteroecker2011}

\bibitem[{O'Meara(2010)}]{O'Meara2010}
O'Meara, B.~C. 2010. {New heuristic methods for joint species delimitation and
  species tree inference.} Systematic biology 59:59--73.
\bibAnnoteFile{O'Meara2010}

\bibitem[{Padial et~al.(2010)Padial, Miralles, {De la Riva}, and
  Vences}]{Padial2010}
Padial, J.~M., A.~Miralles, I.~{De la Riva}, and M.~Vences. 2010. {The
  integrative future of taxonomy}. Frontiers in Zoology 7:1--14.
\bibAnnoteFile{Padial2010}

\bibitem[{Parham et~al.(2013)Parham, Papenfuss, Dijk, Wilson, Marte, Schettino,
  and {Brian Simison}}]{Parham2013}
Parham, J.~F., T.~J. Papenfuss, P.~P.~V. Dijk, B.~S. Wilson, C.~Marte, L.~R.
  Schettino, and W.~{Brian Simison}. 2013. {Genetic introgression and
  hybridization in Antillean freshwater turtles (Trachemys) revealed by
  coalescent analyses of mitochondrial and cloned nuclear markers.} Molecular
  phylogenetics and evolution 67:176--87.
\bibAnnoteFile{Parham2013}

\bibitem[{Pfenninger and Schwenk(2007)}]{Pfenninger2007}
Pfenninger, M. and K.~Schwenk. 2007. {Cryptic animal species are homogeneously
  distributed among taxa and biogeographical regions.} BMC evolutionary biology
  7:121.
\bibAnnoteFile{Pfenninger2007}

\bibitem[{Polly(2003)}]{Polly2003}
Polly, P.~D. 2003. {Paleophylogeography of Sorex araneus: molar shape as a
  morphological marker for fossil shrews}. Mammalia 68:233--243.
\bibAnnoteFile{Polly2003}

\bibitem[{Polly(2007)}]{Polly2007a}
Polly, P.~D. 2007. {Phylogeographic differentiation in Sorex araneus:
  morphology in relation to geography and karyotype}. Russian Journal of
  Theriology 6:73--84.
\bibAnnoteFile{Polly2007a}

\bibitem[{Polly et~al.()Polly, Stayton, Dumont, Pierce, Rayfield, and
  Angielczyk}]{Pollyb}
Polly, P.~D., C.~T. Stayton, E.~R. Dumont, S.~E. Pierce, E.~J. Rayfield, and
  K.~D. Angielczyk. ???? {Combining geometric morphometrics and finite element
  analysis with evolutionary modeling: towards a synthesis}. Journal of
  Vertebrate Paleontology .
\bibAnnoteFile{Pollyb}

\bibitem[{Pons et~al.(2006)Pons, Barraclough, Gomez-Zurita, Cardoso, Duran,
  Hazell, Kamoun, Sumlin, and Vogler}]{Pons2006}
Pons, J., T.~Barraclough, J.~Gomez-Zurita, A.~Cardoso, D.~Duran, S.~Hazell,
  S.~Kamoun, W.~Sumlin, and A.~Vogler. 2006. {Sequence-Based Species
  Delimitation for the DNA Taxonomy of Undescribed Insects}. Systematic Biology
  55:595--609.
\bibAnnoteFile{Pons2006}

\bibitem[{{R Core Team}(2013)}]{2013}
{R Core Team}. 2013. R: A Language and Environment for Statistical Computing. R
  Foundation for Statistical Computing Vienna, Austria.
\bibAnnoteFile{2013}

\bibitem[{Rivera(2008)}]{Rivera2008}
Rivera, G. 2008. {Ecomorphological variation in shell shape of the freshwater
  turtle Pseudemys concinna inhabiting different aquatic flow regimes.}
  Integrative and comparative biology 48:769--87.
\bibAnnoteFile{Rivera2008}

\bibitem[{Rivera et~al.(2014)Rivera, Davis, Godwin, and Adams}]{Rivera2014}
Rivera, G., J.~N. Davis, J.~C. Godwin, and D.~C. Adams. 2014. {Repeatability of
  Habitat-Associated Divergence in Shell Shape of Turtles}. Evolutionary
  Biology Pages~29--37.
\bibAnnoteFile{Rivera2014}

\bibitem[{Rivera and Stayton(2011)}]{Rivera2011}
Rivera, G. and C.~T. Stayton. 2011. {Finite element modeling of shell shape in
  the freshwater turtle Pseudemys concinna reveals a trade-off between
  mechanical strength and hydrodynamic efficiency.} Journal of morphology
  272:1192--203.
\bibAnnoteFile{Rivera2011}

\bibitem[{Rohlf(2005)}]{Rohlf2005}
Rohlf, F.~J. 2005. {TpsDig 2.04}.
\bibAnnoteFile{Rohlf2005}

\bibitem[{Schilck-Steiner et~al.(2007)Schilck-Steiner, Seifert, Stauffer,
  Christian, Crozier, and Steiner}]{SchlickSteiner2007}
Schilck-Steiner, B.~C., B.~Seifert, C.~Stauffer, E.~Christian, R.~H. Crozier,
  and F.~M. Steiner. 2007. {Without morphology, cryptic species stay in
  taxonomic crypsis following discovery}. Trends in ecology \& evolution
  22:391--392.
\bibAnnoteFile{SchlickSteiner2007}

\bibitem[{Seeliger(1945)}]{Seeliger1945}
Seeliger, L.~M. 1945. {Variation in the Pacific Mud Turtle}. Copeia
  1945:150--159.
\bibAnnoteFile{Seeliger1945}

\bibitem[{Spinks and Shaffer(2005)}]{Spinks2005}
Spinks, P.~Q. and H.~B. Shaffer. 2005. {Range-wide molecular analysis of the
  western pond turtle (Emys marmorata): cryptic variation, isolation by
  distance, and their conservation implications.} Molecular ecology
  14:2047--64.
\bibAnnoteFile{Spinks2005}

\bibitem[{Spinks and Shaffer(2009)}]{Spinks2009}
Spinks, P.~Q. and H.~B. Shaffer. 2009. {Conflicting mitochondrial and nuclear
  phylogenies for the widely disjunct Emys (Testudines: Emydidae) species
  complex, and what they tell us about biogeography and hybridization.}
  Systematic biology 58:1--20.
\bibAnnoteFile{Spinks2009}

\bibitem[{Spinks et~al.(2014)Spinks, Thomson, and {Bradley
  Shaffer}}]{Spinks2014}
Spinks, P.~Q., R.~C. Thomson, and H.~{Bradley Shaffer}. 2014. {The advantages
  of going large: genome wide SNPs clarify the complex population history and
  systematics of the threatened western pond turtle}. Molecular Ecology
  Pages~n/a--n/a.
\bibAnnoteFile{Spinks2014}

\bibitem[{Spinks et~al.(2010)Spinks, Thomson, and Shaffer}]{Spinks2010}
Spinks, P.~Q., R.~C. Thomson, and H.~B. Shaffer. 2010. {Nuclear gene
  phylogeography reveals the historical legacy of an ancient inland sea on
  lineages of the western pond turtle, Emys marmorata in California.} Molecular
  ecology 19:542--56.
\bibAnnoteFile{Spinks2010}

\bibitem[{Stayton(2011)}]{Stayton2011}
Stayton, C.~T. 2011. {Biomechanics on the half shell: functional performance
  influences patterns of morphological variation in the emydid turtle
  carapace.} Zoology (Jena, Germany) 114:213--23.
\bibAnnoteFile{Stayton2011}

\bibitem[{Stuart et~al.(2006)Stuart, Inger, and Voris}]{Stuart2006}
Stuart, B.~L., R.~F. Inger, and H.~K. Voris. 2006. {High level of cryptic
  species diversity revealed by sympatric lineages of Southeast Asian forest
  frogs.} Biology letters 2:470--4.
\bibAnnoteFile{Stuart2006}

\bibitem[{Sztencel-Jabłonka et~al.(2009)Sztencel-Jabłonka, Jones, and
  BogdanowicZ}]{Sztencel-Jabonka2009}
Sztencel-Jabłonka, A., G.~Jones, and W.~BogdanowicZ. 2009. {Skull Morphology
  of Two Cryptic Bat Species: Pipistrellus pipistrellus and P. pygmaeus — A
  3D Geometric Morphometrics Approach with Landmark Reconstruction}. Acta
  Chiropterologica 11:113--126.
\bibAnnoteFile{Sztencel-Jabonka2009}

\bibitem[{Tibshirani et~al.(2001)Tibshirani, Walther, and
  Hastie}]{Tibshirani2001a}
Tibshirani, R., G.~Walther, and T.~Hastie. 2001. {Estimating the number of
  clusters in a data set via the gap statistic}. Journal of the Royal
  Statistical Society: Series B (Statistical Methodology) 63:411--423.
\bibAnnoteFile{Tibshirani2001a}

\bibitem[{{Van Bocxlaer} and Hunt(2013)}]{VanBocxlaer2013}
{Van Bocxlaer}, B. and G.~Hunt. 2013. {Morphological stasis in an ongoing
  gastropod radiation from Lake Malawi}. Proceedings of the National Academy of
  Sciences .
\bibAnnoteFile{VanBocxlaer2013}

\bibitem[{van~den Brink and Bokma(2011)}]{VandenBrink2011}
van~den Brink, V. and F.~Bokma. 2011. {Morphometric shape analysis using
  learning vector quantization neural networks — an example distinguishing
  two microtine vole species}. Annales Zoologici Fennici 48:359--364.
\bibAnnoteFile{VandenBrink2011}

\bibitem[{Venables and Ripley(2002)}]{Venables2002a}
Venables, W. and B.~D. Ripley. 2002. {Modern applied statistics with S}. 4th
  ed. Springer, New York.
\bibAnnoteFile{Venables2002a}

\bibitem[{Yang and Rannala(2010)}]{Yang2010b}
Yang, Z. and B.~Rannala. 2010. {Bayesian species delimitation using multilocus
  sequence data.} Proceedings of the National Academy of Sciences 107:9264--9.
\bibAnnoteFile{Yang2010b}

\bibitem[{Zelditch et~al.(2004)Zelditch, Swiderski, and Sheets}]{Zelditch2004}
Zelditch, M.~L., D.~L. Swiderski, and H.~D. Sheets. 2004. {Geometric
  morphometrics for biologists: a primer}. Elsevier Academic Press, Amsterdam.
\bibAnnoteFile{Zelditch2004}

\end{thebibliography}

@@

\begin{table}[ht]
  \centering
  \caption{AUC values for the best model of each classification scheme for both the observed (training) data and the generalized (testing) data. Results from all three different supervised learning approaches are shown here. AUC values range between 0.5 and 1. }
%  \begin{tabular}{ l | c c | c c | c c }
%    \hline
%    & \multicolumn{2}{| l |}{random forest} & 
%    \multicolumn{2}{| l |}{multinomial logistic regression} & 
%    \multicolumn{2}{| l}{linear discriminate analysis} \\
%    Scheme & Observed & Generalized & Observed & Generalized & Observed & Geneeralized \\ 
%    \hline
%    \hline
%    Morph 1 & 0.63 & 0.73 & 0.75 & 0.79 & 0.75 & 0.80 \\ 
%    Morph 2 & 0.61 & 0.58 & 0.76 & 0.77 & 0.76 & 0.77 \\ 
%    Mito 1& 0.63 & 0.62 & 0.75 & 0.63 & 0.75 & 0.63 \\ 
%    Mito 2 & 0.77 & 0.67 & 0.80 & 0.64 & 0.80 & 0.63 \\ 
%    Mito 3 & 0.56 & 0.64 & 0.71 & 0.74 & 0.71 & 0.73 \\ 
%    Nuclear & 0.56 & 0.67 & 0.74 & 0.62 & 0.74 & 0.77 \\ 
%    \hline
%  \end{tabular}
  \label{tab:comp}
\end{table}

\begin{table}[ht]
  \centering
  \caption{Results of bootstrap comparisons between the scheme with the highest mean AUC value and all other schemes, defined P(best - other \(>\) 0). An asterix indicates the best scheme. This was done for each of the three modeling techniques included in this study: random forest (RF), multinomial logistic regression (MLR), and linear discriminate analysis (LDA). Probabilities are the percent of comparisons that are greater than the observed difference in means.}
%  \begin{tabular}{ l r r r }
%    \hline
%    Scheme & RF & MLR & LDA \\
%    \hline
%    \hline
%    Morph 1 & \(\ast\) & \(\ast\) & 1 \\
%    Morph 2 & 0.79 & 0.55 & 1 \\
%    Mito 1 & 0.89 & 0.94 & 1 \\ 
%    Mito 2 & 0.82 & 0.57 & \(\ast\) \\ 
%    Mito 3 & 0.79 & 0.69 & 0.73 \\ 
%    Nuclear & 0.79 & 0.96 & 0.96 \\ 
%    \hline
%  \end{tabular}
  \label{tab:gen_tests}
\end{table}
